---
title: "Fix or Calculate Offsets"
author: "Nicole Barker"
date: "Last run: Jan 4, 2018"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

## Script Abstract

## Background


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(
  fig.path = "figures/"
)
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```

### Load pre-processed bird data

* No screening for temporal alignment or distance to edge
* This should be all the bird data received from Trish in November 2017.

``` {r}
birddat1 <- read.csv("data/birddata_preprocess1.csv", header=T)
colnames(birddat1)
colnames(birddat1)[which(colnames(birddat1) == "PCODE_derived")] <- "PCODE"
colnames(birddat1)[which(colnames(birddat1) == "SS_derived")] <- "SS"
```

**Check for Missing Data**

``` {r}
kable(as.data.frame(do.call(rbind,lapply(birddat1, function(x) {sum(is.na(x))})))) # check for missing data
```



### Combine with Peter's "total bird" offsets

**First, a quality check**

* How many of my kept PKEYs do not have total bird offsets? 
* Is it worth Peter recalculating offsets for them?

Ran this code previously to cache the most recent offset file I have
`offsets <- read.csv("data/offset-species-2016-12-13.csv", header=T)`

`cache("offsets")`

``` {r}
load("cache/offsets.RData")

colnames(offsets)
head(offsets$X)

offsetpkeys <-offsets$X
bcpkeys <- unique(birddat1$PKEY)

load("cache/TOTA_offsets-v3_2016-12-01.Rdata")
tota <- OFF; rm(OFF)
tota <- as.data.frame(tota)
tota$PKEY <- row.names(tota)
colnames(tota)
totapkeys <- tota$PKEY

```

##### Which PKEYs from BC are not in Peter's offsets? 

* `sum(bcpkeys %in% offsetpkeys)` `r sum(bcpkeys %in% offsetpkeys)` of `r length(bcpkeys)` BC Pkeys have offsets
* This suggests that `r length(bcpkeys) - sum(bcpkeys %in% offsetpkeys)` BC PKEYS do __not__ have offsets


##### What about the total species offsets? 
* `sum(bcpkeys %in% totapkeys)` `r sum(bcpkeys %in% totapkeys)` of `r length(bcpkeys)` BC Pkeys have offsets
* This suggests that `r length(bcpkeys) - sum(bcpkeys %in% totapkeys)` BC PKEYS do __not__ have offsets

##### Which PKEYs don't have offsets? 

``` {r}
df.offsettest <- data.frame(OFFSET.indivspp="yes", PKEY=offsets$X)
test <- merge(birddat1, df.offsettest, by="PKEY", all.x=T)
df.totatest <- data.frame(OFFSET.tota="yes", PKEY=totapkeys)
test <- merge(test, df.totatest, by="PKEY", all.x=T)

test$BothMissing <- (test$OFFSET.indivspp == "yes") & (test$OFFSET.tota == "yes")

nooffset <- test[is.na(test$BothMissing),]
write.table(nooffset, "output/BCdat_nooffset_2018.01.04.csv", sep=",", row.names = F, col.names = T)

nooffset$PCODE <- unlist(lapply(strsplit(as.character(nooffset$SS), ":", fixed=T), function(x) {x[1]}))

kable(rbind(head(nooffset), tail(nooffset)),)

length(unique(nooffset$PKEY))

unique(nooffset$PCODE)
```

##### Examine covariates Trish sent to Peter to understand why we're missing offsets for so many PKEYS

``` {r}
offcovar <- read.csv("data/covariatesforoffsetMay2017.csv", header=T)
colnames(offcovar)
unique(offcovar$PCODE)

offtz <- read.csv("data/covariatesforoffsetMay2017Timezone.csv", header=T)
colnames(offtz)
unique(offtz$PCODE)

offcovar.uniquevalues <- as.data.frame(do.call(rbind,lapply(offcovar, function(x) { length(unique(x))})))
colnames(offcovar.uniquevalues) <- c("Num.Unique.Values")
kable(offcovar.uniquevalues)

```

**Initial Notes**

* Looks like all PCODEs (projects) missing offsets are in Trish's covariate data
* This suggests that there was an error in creating the offsets -- somehow these projects were excluded. 
* I'll check for a couple of individual SSs rather than PCODES

``` {r}
randomss.nooff <- sample(nooffset$SS, 10)
offcovar[offcovar$SS %in% randomss.nooff,]
offtz[offtz$SS %in% randomss.nooff,]
```

* Only `r sum(offtz$SS %in% randomss.nooff)` of the SS without offsets are in the covariate datafile
* Somehow some of the SS from the BC data are missing covariates. 
    * Which ones? 
    
    
##### STEPS  

* Identify which PKEYs are in the new PCODES
* How many of them have tota offsets from Peter's May update? 
* How many of them don't have tota offsets from Peter's May update? 
* How many of them have covariates from the file sent to Peter in May? 
* How many of them don't have covariates in the file sent to Peter in May? 

**PKEYS in PCODES that have missing offsets**

``` {r}
new.pcodes <- c(unique(as.character(nooffset$PCODE)), unique(as.character(offcovar$PCODE))) #pcodes with missing offsets
new.pcodes <- data.frame(PCODE=new.pcodes[!duplicated(new.pcodes)]) #pcodes with missing offsets
new.pkeys <- merge(new.pcodes, birddat1, by="PCODE", all.x=T) #PKEYS in those PCODES
nrow(new.pkeys) #how many PKEYS do we need new offsets for

new.ss <- merge(new.pcodes, birddat1[c("METHOD", "SS", "YYYY", "MaxDuration", "Maxdist", "PCODE")], by="PCODE", all.x=T) #PKEYS in those PCODES
new.ss  <- new.ss[!duplicated(new.ss),]
nrow(new.ss)
```

**SS with tota offsets from May update**

* Offsets are generated on a per-PKEY basis
* But covariates are provided on a per-SS basis, so I need to extract SS from this offset file

``` {r}
pkey.has.offsets <- data.frame(PKEY=unique(tota$PKEY), hastotaoffset="yes")
pkey.has.offsets$SS <- unlist(lapply(strsplit(as.character(pkey.has.offsets$PKEY), ":", fixed=T), function(x) {paste(x[1:3], collapse=":")}))
ss.has.offsets <- pkey.has.offsets[c("SS", "hastotaoffset")]
ss.has.offsets <- ss.has.offsets[!duplicated(ss.has.offsets),]

new.ss <- merge(new.ss, ss.has.offsets, by="SS", all.x=T)

new.ss$hastotaoffset <- as.character(new.ss$hastotaoffset)
new.ss$hastotaoffset[is.na(new.ss$hastotaoffset)] <- "no"
kable(rbind(head(new.ss), tail(new.ss)))

sum(new.ss$hastotaoffset == "yes")
sum(new.ss$hastotaoffset == "no")
nrow(new.ss)
```

**SS in covariate file Trish sent to Peter in May**

``` {r}
ss.has.offset.covars <- data.frame(SS=unique(offcovar$SS), hascovar="yes")
ss.has.offset.covars <- ss.has.offset.covars[!duplicated(ss.has.offset.covars),]

new.ss <- merge(new.ss, ss.has.offset.covars, by="SS", all.x=T)

new.ss$hascovar <- as.character(new.ss$hascovar)
new.ss$hascovar[is.na(new.ss$hascovar)] <- "no"
kable(rbind(head(new.ss), tail(new.ss)))

sum(new.ss$hascovar == "yes")
sum(new.ss$hascovar == "no")
nrow(new.ss)
```

``` {r}
ss.has.offset.tzcovars <- data.frame(SS=unique(offtz$SS), hastzcovar="yes")
ss.has.offset.tzcovars <- ss.has.offset.tzcovars[!duplicated(ss.has.offset.tzcovars),]

new.ss <- merge(new.ss, ss.has.offset.tzcovars, by="SS", all.x=T)

new.ss$hastzcovar <- as.character(new.ss$hastzcovar)
new.ss$hastzcovar[is.na(new.ss$hastzcovar)] <- "no"
kable(rbind(head(new.ss), tail(new.ss)))

sum(new.ss$hastzcovar == "yes")
sum(new.ss$hascovar == "no")
nrow(new.ss)
```

**NOTES**

* Makes sense that we would be missing covariates for some of the offsets we're missing 
* Not sure why we are missing offsets for some we have covariates for
* How can we possibly have offsets for some we don't have covariates for?

**I suspect a problem in the naming, in switching from SS to PKEY or vice versa**

* But really I don't have time to dig into the details and understand the mistake. How do we fix it? 


## Can I create table with all covariates needed to generate offsets? At PKEY level? 

* Covariates needed for offsets: Nalsms, percent tree cover, lat/lon, and 1st day of spring (this latter is optional), tz is also helpful

* Extract point count method details from birddat table - at SS level
* Extract covariates from Trish's covariate table - also at SS level
* Combine them
* Merge with birddat table[SS, PKEY] to get all PKEYS I need offsets for
* Merge with current tota offsets to demonstrate which PKEYs currently have offsets and which are missing

``` {r}
covar.ss <- offtz[c("SS", "tree", "NALCMS05", "TZID", "FID_tz_wor", "POINT_X", "POINT_Y")]
covar.ss <- covar.ss[!duplicated(covar.ss),]

birddat2 <- birddat1[c("SS", "METHOD", "YYYY", "MaxDuration", "Maxdist", "PCODE")]
birddat2 <- birddat2[!duplicated(birddat1),]

new.dat.for.offsets <- merge(birddat2, covar.ss, by="SS", all.x = T)

new.dat.for.offsets <- new.dat.for.offsets[!duplicated(new.dat.for.offsets),]

kable(new.dat.for.offsets[sample(nrow(new.dat.for.offsets), 15),], row.names=F)

```

**Check for Missing Data**

``` {r}
kable(as.data.frame(do.call(rbind, lapply(new.dat.for.offsets, function(x) {sum(is.na(x))}))), row.names=F)

missingcovar <- new.dat.for.offsets[is.na(new.dat.for.offsets$tree),] #use 'tree' as an example covariate

(pcodes.missingcovar <- unique(missingcovar$PCODE))
ss.missingcovar <- unique(missingcovar$SS)
length(ss.missingcovar)
```

##### Combine with PKEY by SS to make a table of covariates required to calculate offsets by PKEY

``` {r}
pkey.ss <- birddat1[c("PKEY", "SS")]
pkey.ss <- pkey.ss[!duplicated(pkey.ss),]

new.dat.for.offsets2 <- merge(new.dat.for.offsets, pkey.ss, by="SS", all.x = T)
new.dat.for.offsets2 <- merge(new.dat.for.offsets2, pkey.has.offsets[c("PKEY", "hastotaoffset")], by="PKEY", all.x=T)

kable(new.dat.for.offsets2[sample(nrow(new.dat.for.offsets2), 15),], row.names=F)
```

**Check for missing data**

``` {r}

as.data.frame(do.call(rbind, lapply(new.dat.for.offsets2, function(x) {sum(is.na(x))})))

```

```{r}
write.table(new.dat.for.offsets2, "output/new.dat.for.offsets.csv", col.names=T, row.names=F, sep=",")
```

#### Re-calculate offsets since we're missing a bunch 

Code supplied by Peter Solymos on Jan 2, 2018. 

* Note that this method uses only max distance and max duration, so it only corrects for methodology. Not location- or time-specific covariates. 

``` {r, eval=T}
require(QPAD)
x <- birddat1
str(x)
table(x$MaxDuration, useNA="a")
table(x$Maxdist, useNA="a")
levels(x$Maxdist)[levels(x$Maxdist) == "Unlimited"] <- "Inf"
x$MaxDistance <- as.numeric(as.character(x$Maxdist)) / 100
table(x$MaxDistance, useNA="a")

# numbers taken from estimates
phi_tota <- exp(0.2773876)
tau_tota <- exp(0.8422723)
x$p <- sra_fun(x$MaxDuration, phi_tota)
x$A <- ifelse(is.finite(x$MaxDistance), x$MaxDistance^2*pi, tau_tota^2*phi_tota)
x$q <- ifelse(is.finite(x$MaxDistance), edr_fun(x$MaxDistance, tau_tota), 1)
summary(x$p)
summary(x$A)
summary(x$q)

x$Exposure <- with(x, p * A * q)
x$Offset <- log(x$Exposure)
summary(x$Offset)
length(x$Offset)

sum(is.na(x$Offset)) # how many PKEYs don't have offsets
```


**If applicable, recombine the calculated offsets with Peter's previously calculated offsets**

As of Jan 3, 2017. I re-calculated all offsets using the simplified approach (above), so this chunk isn't necessary.

``` {r eval=F}
row.names(x) <- x$PKEY
colnames(x)[which(colnames(x) == "Offset")] <- "TOTA"
x <- x[c("TOTA", "PKEY")]
tota2 <- rbind(tota, x)
sum(duplicated(tota2$PKEY)) #none are duplicated!

birddat5 <- merge(birddat1, tota2, by="PKEY", all.x=T)
sum(is.na(birddat5$TOTA)) # how many PKEYs don't have offsets
```

### Cache pre-processed bird data

``` {r}
birddata_preprocess2_withOffset <- x
cache("birddata_preprocess2_withOffset")
```