---
title: "Preprocess Full Forsite VRI Files"
author: "Nicole Barker"
date: "Last run: Jan 8, 2018"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---
## Script Abstract
Initial inspection of the attribute table exported from the Forsite-modified version of the VRI dataset. Identifies which field is a unique identifier. Saves by caching as RData to the cache folder. 

## Background

**FILES**

1. FORSITE_all_regions__Nicoles_VRI_fields_fixednames.csv

In this script, I: 

* Inspect for duplicates
* Find a unique stand identifier to link back to the Shapefile (for eventual prediction)

Note that CanfortflAttributeTable.txt is processed in another file. 


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```


## migratory_bird_BAMused.txt

### Filter out some columns I know I won't need

``` {r, eval=T}
if ("vri_forsite_mostfields.RData" %in% dir("cache") == TRUE) { #conditional statements to automate loading and caching 
  load("cache/vri_forsite_mostfields.RData")                    #on existence of cached object.
  } 

if ("vri_forsite_mostfields.RData" %in% dir("cache") == FALSE) { #conditional statements to automate loading and caching 
  if ("vri_forsite_allfields.RData" %in% dir("cache") == TRUE) { #conditional statements to automate loading and caching  
    load("cache/vri_forsite_allfields.RData")                    #on existence of cached object.
    }
    if ("vri_forsite_allfields.RData" %in% dir("cache") == FALSE) { #conditional statements to automate loading and caching 
      vri_forsite_allfields <- read.csv("data/migratory_bird_BAMused.txt", header=T)
      cache("vri_forsite_allfields")
      }
    
    cols_to_remove <- colnames(vri_forsite_allfields)[grep("DEAD_VOL", colnames(vri_forsite_allfields))]
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("LIVE_VOL", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("BIOMASS", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("LINE_", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("LABEL_", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("STAND_VOL", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("CRUISE", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("LABEL", colnames(vri_forsite_allfields))])
    cols_to_remove <- c(cols_to_remove, colnames(vri_forsite_allfields)[grep("QUAD", colnames(vri_forsite_allfields))])
    
    vri_forsite_mostfields <- vri_forsite_allfields[-which(colnames(vri_forsite_allfields) %in% cols_to_remove)]
    cache("vri_forsite_mostfields")
}
```

``` {r}
kable(data.frame(ColNames=colnames(vri_forsite_mostfields)))
vri_forsite <- vri_forsite_mostfields;rm(vri_forsite_mostfields)
```

#### Initial exploration to understand the format of the shapefile attribute table

``` {r}
tmp <- as.data.frame(do.call(rbind, lapply(vri_forsite, function(x) {length(unique(x))})))
tmp[which.max(tmp$V1),] == nrow(vri_forsite) # check to make sure there is a unique ID for stands
rownames(tmp)[which.max(tmp$V1)]
```

* Number of unique objects in the polygon attribute table (OBJECT_ID): `r length(unique(vri_forsite$OBJECTID))`
* Number of unique polygons in the polygon attribute table (FEATURE_ID): `r length(unique(vri_forsite$FEATURE_ID))`
* Number of rows in the polygon attribute table: `r nrow(vri_forsite)`
* Number of unique forest stand types in the polygon attribute table (ForID): `r length(unique(vri_forsite$ForID))` 
* Number of possible unique forest stand types based on combination of 4 attributes: `r length(unique(mastermatrix$ForID))`

**Conclusions:** 

* both OBJECTID and FEATURE_ID are unique identifiers for rows (polygons) in this file.
* No duplicates because same number of rows as unique polygon IDs. 

## Add in polygon area

In a previous analysis, I calculated the area of each polygon using the calculate geometry command in ArcMap. 
Here I'm merging polygon area with the remaining VRI attributes in this file. 

``` {r}
load("cache/mig_birds_Layer_polyarea_forstandatts.RData")
vri_forsite2 <- merge(mig_birds_Layer_polyarea_forstandatts[c("FEATURE_ID", "polygon_area_cal")], vri_forsite, by="FEATURE_ID", all=T)
nrow(vri_forsite) == nrow(vri_forsite2)
vri_forsite <- vri_forsite2; rm(vri_forsite2); gc()
```

#### Add source field and a unique ID field

``` {r}
vri_forsite$VRI_Layer <- "Forsite_VRI"
vri_forsite$StandID <- paste(vri_forsite$VRI_Layer, vri_forsite$OBJECTID, sep=".")
length(unique(vri_forsite$StandID))
nrow(vri_forsite)
```

#### Save file

``` {r}
cache("vri_forsite")
```
