---
title: "Preprocess Canfor TFL VRI File"
author: "Nicole Barker"
date: "Last run: Dec 20, 2017"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

## Background

**FILES**

1. CanfortflAttributeTable.txt

In this script, I: 

* Inspect for duplicates

Note that FORSITE_all_regions__Nicoles_VRI_fields_fixednames.csv is processed in another file. 


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```

## canfortflAttributeTable.txt

``` {r}
vri_canfortfl <- read.csv("data/canfortflAttributeTable.txt", header=T)
```

#### Initial exploration to understand the format of the shapefile attribute table

* Number of unique objects in the polygon attribute table (OBJECTID): `r length(unique(vri_canfortfl$OBJECTID))`
* Length of the attribute table (nrows): `r nrow(vri_canfortfl)`

#### Save file

``` {r}
cache("vri_canfortfl")
```
