---
title: "Preprocess Full Forsite VRI Files"
author: "Nicole Barker"
date: "Last run: Dec 20, 2017"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

## Background

**FILES**

1. FORSITE_all_regions__Nicoles_VRI_fields_fixednames.csv

In this script, I: 

* Inspect for duplicates

Note that CanfortflAttributeTable.txt is processed in another file. 


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```


## FORSITE_all_regions__Nicoles_VRI_fields_fixednames.csv

This is an export of the attribute table from "M:\DataStuff\SpatialData\VegetationResourceInventory\2016.10_Download\FORSITE_all_regions_all_vri_fields.gdb", dropped to the fields I wanted. 

The "fixednames" is because the table exported with incomplete field names. This was fixed in a previous iteration of analyses, so I don't need to repeat it now. 

``` {r}
vri_forsite <- read.csv("data/FORSITE_all_regions__Nicoles_VRI_fields_fixednames.csv", header=T)
kable(data.frame(ColNames=colnames(vri_forsite)))
```

#### Initial exploration to understand the format of the shapefile attribute table

* Number of unique objects in the polygon attribute table (OBJECT_ID): `r length(unique(vri_forsite$OBJECTID))`
* Number of unique polygons in the polygon attribute table (FEATURE_ID): `r length(unique(vri_forsite$FEATURE_ID))`
* Number of rows in the polygon attribute table: `r nrow(vri_forsite)`
* Number of unique forest stand types in the polygon attribute table (ForID): `r length(unique(vri_forsite$ForID))` 
* Number of possible unique forest stand types based on combination of 4 attributes: `r length(unique(mastermatrix$ForID))`

**Conclusions:** 

* both OBJECTID and FEATURE_ID are unique identifiers for rows (polygons) in this file.
* No duplicates because same number of rows as unique polygon IDs. 

## Add in polygon area

In a previous analysis, I calculated the area of each polygon using the calculate geometry command in ArcMap. 
Here I'm merging polygon area with the remaining VRI attributes in this file. 

``` {r}
load("cache/mig_birds_Layer_polyarea_forstandatts.RData")
vri_forsite <- merge(mig_birds_Layer_polyarea_forstandatts[c("FEATURE_ID", "polygon_area_cal")], vri_forsite, by="FEATURE_ID", all=T)
```

#### Save file

``` {r}
cache("vri_forsite")
```
