---
title: "Merge Forsite and Canfor TFL VRI files together"
author: "Nicole Barker"
date: "Last run: Dec 22, 2017"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

## Script Abstract
Identify which columns are needed from both VRI files (Canfor and Forsite). Add blank columns to Canfor file so it can be combind with the Forsite file. Combine the two files together and modify the factors to have the correct levels. 

## Background

Forsite's VRI file has blank attributes for tree farm licences (TFLs) that Canfor has data for. Canfor gave us their VRI data for those forest stands. I separately quality-checked and processed the two VRI datasets and now it's time to combine them, after doing a bit more processing. 

**FILES**

1. vri_forsite_zSpGroup_Rankfixed.RData
2. vri_canfortfl.RData

In this script, I: 

* Look for overlapping columns between two VRI files
* create blank columns in Canfor table where I need them
* Combine two VRI files together
* Convert date columns to date format 
* Pull year out from date
* Classify stands to a stand age class if not already done
* Classify stands to a stange height class if not already done
* Create multiple files for future use 
    * Forsite VRI: for analyses using Kari/Forsite's 4 derived attributes (will exclude Canfor TFL); and
    * All VRI: for density models (using variables from VRI directly, not derived)


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```

### Look for overlapping columns between two VRI files

``` {r}
load("cache/vri_forsite_zSpGroup_Rankfixed.RData")
vri_forsite <- vri_forsite_zSpGroup_Rankfixed; rm(vri_forsite_zSpGroup_Rankfixed)

forsite_colnames <- data.frame(ColName=colnames(vri_forsite), In_ForsiteVRI=as.character("Yes"))

load("cache/vri_canfortfl.RData")
canfor_colnames <- data.frame(ColName=colnames(vri_canfortfl), In_vri_canfortfl=as.character("Yes"))
```

``` {r}
all_colnames <- merge(canfor_colnames, forsite_colnames, by="ColName", all=T)

all_colnames$In_vri_canfortfl <- as.character(all_colnames$In_vri_canfortfl)
all_colnames$In_ForsiteVRI <- as.character(all_colnames$In_ForsiteVRI)

all_colnames$In_vri_canfortfl[is.na(all_colnames$In_vri_canfortfl)] <- ""
all_colnames$In_ForsiteVRI[is.na(all_colnames$In_ForsiteVRI)] <- ""

kable(all_colnames, row.names=F)
```

#### Add blank columns for those I need from the Forsite layer (i.e. derived fields)

``` {r}
vri_canfortfl$BecZone <- NA                 
vri_canfortfl$SpeciesGroup <- NA
vri_canfortfl$Age <- NA
vri_canfortfl$Height <- NA
vri_canfortfl$Rank <- NA
vri_canfortfl$FD_PERCENT <- NA
vri_canfortfl$PL_PERCENT <- NA
vri_canfortfl$AT_PERCENT <- NA
vri_canfortfl$Decid_PERCENT <- NA
vri_canfortfl$polygon_area_cal <- NA
vri_canfortfl$FEATURE_ID <- NA
vri_forsite$OBJECTID <- NA

canfor_colnames <- data.frame(ColName=colnames(vri_canfortfl), In_vri_canfortfl=as.character("Yes"))

all_colnames <- merge(canfor_colnames, forsite_colnames, by="ColName", all=T)

all_colnames$In_vri_canfortfl <- as.character(all_colnames$In_vri_canfortfl)
all_colnames$In_ForsiteVRI <- as.character(all_colnames$In_ForsiteVRI)

all_colnames$In_vri_canfortfl[is.na(all_colnames$In_vri_canfortfl)] <- ""
all_colnames$In_ForsiteVRI[is.na(all_colnames$In_ForsiteVRI)] <- ""

kable(all_colnames, row.names=F)
```

## Merge two VRI files together

#### Subset for common set of columns
``` {r}
commoncols <- all_colnames[all_colnames$In_vri_canfortfl == "Yes" & all_colnames$In_ForsiteVRI == "Yes",]$ColName

vri_canfortfl2 <- vri_canfortfl[which(colnames(vri_canfortfl) %in% commoncols)]
vri_canfortfl2$VRI_Layer <- "Canfor_TFL"
vri_forsite2 <- vri_forsite[which(colnames(vri_forsite) %in% commoncols)]
vri_forsite2$VRI_Layer <- "Forsite_VRI"
```

#### rbind together

``` {r}
vri_both <- rbind(vri_canfortfl2, vri_forsite2)
```

## More pre-processing

### Date pre-processing

``` {r, warning=F, message=F}
F.convertdate.extractyear <- function(x){
  dates <- as.Date(unlist(lapply(strsplit(as.character(x), " "), function(y) {y[1]})), format="%m/%d/%Y")
  years <- format(dates,'%Y')
}

vri_both$InputYear <- F.convertdate.extractyear(vri_both$INPUT_DATE)
vri_both$ProjectedYear <- F.convertdate.extractyear(vri_both$PROJECTED_DATE)
vri_both$InterpretYear <- F.convertdate.extractyear(vri_both$INTERPRETATION_DATE)
vri_both$ReferenceYear <- F.convertdate.extractyear(vri_both$REFERENCE_DATE)
vri_both$AttributionYear <- F.convertdate.extractyear(vri_both$ATTRIBUTION_BASE_DATE)
vri_both$DistYear_Nonlog <- F.convertdate.extractyear(vri_both$EARLIEST_NONLOGGING_DIST_DATE)
vri_both$DistYear_Log <- F.convertdate.extractyear(vri_both$HARVEST_DATE)
vri_both$DisturbanceYear <- apply(vri_both[c("DistYear_Nonlog", "DistYear_Log")], MARGIN=1, FUN=function(x) {max(x, na.rm=T)}) 
```

### Stand Age classification

``` {r}
ageclassbreaks <- c(0, 3, 30, 80, 120, 250, 1000)
vri_both$AgeClass_calc <- cut(vri_both$PROJ_AGE_1, breaks = ageclassbreaks, include.lowest = TRUE, labels = c("0-3 yrs", "3-30 yrs", "31-80 yrs", "81-120 yrs", "121-250 yrs", ">250 yrs"))

kable(rbind(head(vri_both[c("PROJ_AGE_1", "AgeClass_calc", "VRI_Layer")]),
            tail(vri_both[c("PROJ_AGE_1", "AgeClass_calc", "VRI_Layer")])))
```

### Stand Height classification

``` {r}
heightclassbreaks <- c(0, 10.5, 19.5, 28.5, 100)
vri_both$HeightClass_calc <- cut(vri_both$PROJ_HEIGHT_1, breaks = heightclassbreaks, include.lowest = TRUE, labels = c("<10.5 m", "10.5-19.4 m", "19.5-28.4 m", "> 28.5 m"))

kable(rbind(head(vri_both[c("PROJ_HEIGHT_1", "HeightClass_calc", "VRI_Layer")]),
            tail(vri_both[c("PROJ_HEIGHT_1", "HeightClass_calc", "VRI_Layer")])))
```

## Save files

``` {r, eval=T}
cache("vri_both")

vri_canfortfl_postclean <- vri_both[vri_both$VRI_Layer == "Canfor_TFL",]
cache("vri_canfortfl_postclean")

vri_forsite_postclean <- vri_both[vri_both$VRI_Layer == "Forsite_VRI",]
cache("vri_forsite_postclean")
```


