---
title: "Ranks vs. Bird Density"
author: "Nicole Barker"
date: "Last updated: Jan 2, 2018"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

##Script Abstract


# Question: How well do ranks predict observed bird density?

## Approach 1. Descriptive plots of density vs. ranks

**Steps **

* Calculate the average (method/detectability-corrected) density for each forest stand type (ForID) that has at least one point count survey
* To get the average per stand type, I basically build a mean-model (no covariates), which also corrects for method via offsets. 
* Poisson GLM: Count + offset ~ ForID
    * Forest stand type as predictor
    * Count as response variable
    * TOTA offset as offset 
* Data subset: Those stand types that were sampled by at least one bird point count survey.



## Approach 2. Quantitatively analyze relationship between density (count with offset) and Rank


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(
  fig.path = "figures/"
)
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```


## Load data

``` {r}
load("cache/bird_vri_dat_ready.RData")
colnames(bird_vri_dat_ready)
tota <- bird_vri_dat_ready
colnames(tota)[which(colnames(tota) == "TOTA")] <- "Offset"
```

### Filter out some data I want to exclude from ranks vs. bird density analysis

**Stands that have been disturbed by insects**

``` {r}
tota.qs <- tota
tota$EARLIEST_NONLOGGING_DIST_TYPE <- gsub(" *$", "", tota$EARLIEST_NONLOGGING_DIST_TYPE) # remove trailing white space
nrow(tota[tota$EARLIEST_NONLOGGING_DIST_TYPE == "IBM",])
tota <- subset(tota, tota$EARLIEST_NONLOGGING_DIST_TYPE != "IBM")
```

**Surveys with large numbers of birds**

``` {r}
summary(tota$sumABUND)
nrow(tota[tota$sumABUND > 100,])
tota <- subset(tota, tota$sumABUND < 100)
```


**Include only one survey per sampling location (SS)**

A rigorous approach would be repeating model building with multiiple random sub-samples. However, I'm going to extract a single year of data from each SS that has multiple years of data. 

I'll choose the specific year based on its proximity to the VRI reference year. This is just a matter of sorting based on Difference and keeping the first one. 

``` {r}
tota <- tota[order(tota$SS, tota$Difference_Bird_Photo),]

dupSS <- tota$SS[duplicated(tota$SS)] # the duplicated SSs

tota.nodup <- tota[!tota$SS %in% dupSS,]
tota.dupl <- tota[tota$SS %in% dupSS,]

kable(tota.dupl[1:25,c("SS", "PKEY", "Difference_Bird_Photo", "Age")])

tota.dupl <- tota.dupl[!duplicated(tota$SS),] # cuts out all subsequently duplicated rows while keeping the first for each SS

tota <- rbind(tota.nodup, tota.dupl)
colnames(tota)[which(colnames(tota) == "sumABUND")] <- "Count"
```


## Approach 1. Descriptive plots of density vs. ranks

### Calculating average density per forest stand type 

Forest stand type is a factor with `r length(unique(tota$ForID))` levels

``` {r fit.glm.forid, eval=T, echo=T}
if ("fit.dens.forid_tot.RData" %in% dir("cache") == TRUE) { #conditional statements to automate loading and caching 
  load("cache/fit.dens.forid_tot.RData")                    #on existence of cached model object.
}

if(!exists("fit.dens.forid_tot")) {
  fit.dens.forid_tot <- glm(Count ~ ForID-1, family=poisson, data=tota, offset=Offset)
  cache("fit.dens.forid_tot")
  }
```

##### Calculate the number of birds in each forest stand type

Use 'predict' on the simple model above to attribute the mean density value for each forest stand type, after correcting for survey method. 

Only predicts to the forest stands we sampled. `r length(unique(tota$ForID))` unique stand types.

Basically just takes the mean of all samples within a forest stand type and adjusts by the offset. No covariates in the above model.

``` {r glm.forid.pred, eval=T}
tota$Pred_fit.dens.forid_tot <- predict(fit.dens.forid_tot, newdata=tota, type="response")
```

##### Double-check that these predictions make sense. 
Expectation: All point counts within the same forest stand type should have the same expected density, if I understand this analysis correctly. 

``` {r}
testset <- tota[tota$ForID %in% "SBS_SBPS.SwB_Decid.31_80.2",]
testset <- testset[order(testset$Pred_fit.dens.forid_tot, decreasing = F),]
testset$expOffset <- exp(testset$Offset)
testset$PredDensity <- testset$Pred_fit.dens.forid_tot / exp(testset$Offset)
kable(testset[1:10,c("ForID", "PKEY", "Pred_fit.dens.forid_tot", "expOffset", "Offset", "PredDensity")])
```























































### GLM of Density ~ Rank

##### Treat forest stand rank (1-6) as continuous

``` {r}
tota2$Rank <- as.numeric(as.character(tota2$Rank))
tota3 <- tota2[!is.na(tota2$Rank),]
tota3 <- tota3[!is.na(tota3$Offset),]
tota3 <- tota3[!is.na(tota3$sumABUND),]
```

``` {r eval=T, echo=T}
fit1 <- glm(sumABUND ~ Rank, family=poisson, data=tota3, offset=Offset)
summary(fit1)
par(mfrow=c(2,2))
plot(fit1)
par(mfrow=c(1,1))
tota3$pred_fit1 <- predict(fit1, type="response")
plot(tota3$pred_fit1 ~ tota3$sumABUND, xlab="Observed Abundance", ylab="Predicted Abundance")
```

#### Repeat the analysis separately for each Bec Zone, to account for geographic variation

``` {r glm.dens.rank.bybec, echo=T}
tota.becsplit <- split(tota, tota$BecZone)
becglms <- lapply(tota.becsplit, function(x) {
  fitx <- glm(sumABUND ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(becglms), function(i) {
  data.frame(
    BecZone=names(becglms[i]), 
    Intercept=summary(becglms[[i]])$coefficients["(Intercept)","Estimate"], 
    Rankvs.Density=summary(becglms[[i]])$coefficients["Rank","Estimate"], 
    RankSE=summary(becglms[[i]])$coefficients["Rank","Std. Error"])
}))

kable(sumresults, row.names=F)
rm(sumresults)
```

#### Repeat the analysis separately for each species group, to account for tree productivity differences

Note that I removed a couple of species groups because 

``` {r glm.dens.rank.byspp, eval=T, echo=T}
tota.spsplit <- split(tota, tota$SpeciesGroup)
spgrpglms <- lapply(tota.spsplit[c(1:11,13:16, 18:21)], function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(spgrpglms), function(i) {
  data.frame(BEC=names(spgrpglms[i]), Intercept=summary(spgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(spgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(spgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

kable(sumresults, row.names=F)
ggplot(sumresults, aes(x=BEC, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1)

rm(sumresults)
```

#### Repeat the analysis separately for each combination of species group and bec, to account for tree productivity differences

``` {r glm.dens.rank.bybec.byspp, eval=T, echo=T}
tota.sub <- tota[!tota$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota$becspp <- paste(tota$BEC, tota$SpeciesGroup, sep=".")
totas <- split(tota, tota$becspp)
totas <- totas[-c(1,  5,  7, 21, 25, 27, 39, 43, 63, 80, 81, 82)]

becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- lapply(1:length(becspgrpglms), function(i) {
  summary(becspgrpglms[[i]])$coefficients
})

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults <- sumresults[order(sumresults$Rankvs.Density, decreasing=T),]

kable(sumresults, row.names=F)
kable(head(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults)[1] <- "BEC.Species"
sumresults$BEC <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[2]}))

ggplot(sumresults, aes(x=SpeciesGroup, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)


```


**Conclusion:** performance of rank for density dpeends on combination of bec and sp group.

``` {r}
reps <- read.csv("data/Output/BirdSamplingStandTypes-cleaned.2017.04.04.csv", header=T)
becspcountSS <- aggregate(reps$TotalCountSS, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becspcountSS)[ncol(becspcountSS)] <- "CountSS"
becspcountSS$BEC.Species <- paste(becspcountSS$BEC, becspcountSS$SpeciesGroup, sep=".") 

becsparea <- aggregate(reps$ForestArea, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becsparea)[ncol(becsparea)] <- "ForestArea"
becsparea$BEC.Species <- paste(becsparea$BEC, becsparea$SpeciesGroup, sep=".") 

becsp <- merge(becspcountSS, becsparea[c("BEC.Species", "ForestArea")], by="BEC.Species")

sumresults2 <- merge(sumresults, becsp[c("BEC.Species", "ForestArea", "CountSS")], by="BEC.Species", all.x=T)

sumresults2 <- sumresults2[order(sumresults2$CountSS, decreasing=T),]
kable(sumresults2, row.names=F)
```

#### Repeat the analysis including a Rank^2 term. Will maybe give us an idea of why the ranks don't correlate with density.

``` {r glm.dens.rank.rank2.bybec.byspp, eval=F, echo=T}
tota.sub <- tota[!tota$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota$becspp <- paste(tota$BEC, tota$SpeciesGroup, sep=".")
tota$Rank2 <- tota$Rank^2
totas <- split(tota, tota$becspp)


becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank + Rank*Rank, family=poisson, data=x, offset=Offset)
})

removethese <- which(unlist (lapply(1:length(becspgrpglms), function(i) {nrow(summary(becspgrpglms[[i]])$coefficients)}))<3)

totas <- totas[-removethese]

becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank + Rank*Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- lapply(1:length(becspgrpglms), function(i) {
  summary(becspgrpglms[[i]])$coefficients
})

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"],
RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"],
Rank2vs.Density=summary(becspgrpglms[[i]])$coefficients["Rank2","Estimate"],
Rank2SE=summary(becspgrpglms[[i]]$coefficients["Rank2", "Std. Error"]))
}))

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults <- sumresults[order(sumresults$Rankvs.Density, decreasing=T),]

kable(sumresults, row.names=F)
kable(head(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults)[1] <- "BEC.Species"
sumresults$BEC <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[2]}))

ggplot(sumresults, aes(x=SpeciesGroup, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)
```


**Conclusion:** performance of rank for density dpeends on combination of bec and sp group.

``` {r}
reps <- read.csv("data/Output/BirdSamplingStandTypes-cleaned.2017.04.04.csv", header=T)
becspcountSS <- aggregate(reps$TotalCountSS, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becspcountSS)[ncol(becspcountSS)] <- "CountSS"
becspcountSS$BEC.Species <- paste(becspcountSS$BEC, becspcountSS$SpeciesGroup, sep=".") 

becsparea <- aggregate(reps$ForestArea, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becsparea)[ncol(becsparea)] <- "ForestArea"
becsparea$BEC.Species <- paste(becsparea$BEC, becsparea$SpeciesGroup, sep=".") 

becsp <- merge(becspcountSS, becsparea[c("BEC.Species", "ForestArea")], by="BEC.Species")

sumresults2 <- merge(sumresults, becsp[c("BEC.Species", "ForestArea", "CountSS")], by="BEC.Species", all.x=T)

sumresults2 <- sumresults2[order(sumresults2$CountSS, decreasing=T),]
kable(sumresults2, row.names=F)
```


## Try re-running glm after omitting youngest age class
``` {r}
tota_older <- tota[!tota$ForAge %in% "3_30",]

tota.sub <- tota_older[!tota_older$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota.sub$becspp <- paste(tota.sub$BEC, tota.sub$SpeciesGroup, sep=".")
totasubs <- split(tota.sub, tota.sub$becspp)

becspgrpglms <- lapply(totasubs, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

removethese <- which(unlist (lapply(1:length(becspgrpglms), function(i) {nrow(summary(becspgrpglms[[i]])$coefficients)}))==1)

totasubs <- totasubs[-removethese]

becspgrpglms <- lapply(totasubs, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults_noyoung <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults_noyoung <- sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),]

kable(sumresults_noyoung, row.names=F)
kable(head(sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults_noyoung)[1] <- "BEC.Species"
sumresults_noyoung$BEC <- unlist(lapply(strsplit(as.character(sumresults_noyoung$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults_noyoung$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults_noyoung$BEC.Species), ".", fixed=T), function(x){x[2]}))
colnames(sumresults_noyoung)[which(colnames(sumresults_noyoung)=="Rankvs.Density")] <- "Rankvs.Density_noyoung"

ggplot(sumresults_noyoung, aes(x=SpeciesGroup, y=Rankvs.Density_noyoung)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)

test <- merge(sumresults, sumresults_noyoung[c("BEC.Species","Rankvs.Density_noyoung")], by="BEC.Species")
kable(test[c("BEC.Species", "Rankvs.Density","Rankvs.Density_noyoung")], row.names=F)
```






### GLM of Density ~ Rank

Treat rank as a factor to see if those results are different.

``` {r eval=T, echo=T}
tota$Rank <- as.factor(tota$Rank)
fit1 <- glm(Count ~ Rank, family=poisson, data=tota, offset=Offset)
summary(fit1)

```

Looks like density for each rank is slightly higher than density for rank 1, with exception of rank 6...which has roughly the same density as rank1.

#### Repeat separately for each Bec Zone
``` {r eval=T, echo=T}
totas <- split(tota, tota$BEC)
becglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(becglms), function(i) {
  data.frame(BEC=names(becglms[i]), Intercept=summary(becglms[[i]])$coefficients["(Intercept)","Estimate"], Rank2=summary(becglms[[i]])$coefficients["Rank2","Estimate"], Rank3=summary(becglms[[i]])$coefficients["Rank3","Estimate"],
Rank4=summary(becglms[[i]])$coefficients["Rank4","Estimate"],
Rank5=summary(becglms[[i]])$coefficients["Rank5","Estimate"])
}))

kable(sumresults, row.names=F)
```
