---
title: "Ranks Versus Density"
author: "Nicole Barker"
date: "April 6, 2017"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

# Question: How well do ranks predict observed bird density?

Approach: Simple regression of density (count with offset) ~ Rank

``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
```

```{r startup, echo=F, message=FALSE, warning=FALSE}
require(ProjectTemplate)
rm(list=ls())
load.project()
```

```{r load.pointcountloc.data, echo=T, eval=T}
load("cache/TOTAlcount.per.PKEY.withForID.RData")
tota <- TOTAlcount.per.PKEY.withForID
tota$SS <- as.character(tota$SS)
tota <- merge(tota, mastermatrix[c("ForID", "SpeciesGroup", "BEC", "ForAge", "ForHt")], by="ForID")
tota$BEC <- factor(tota$BEC)
tota$Rank <- as.numeric(as.character(tota$Rank))
```

### GLM of Density ~ Rank

Treat forest stand rank (1-6) as continuous

``` {r eval=T, echo=T}
fit1 <- glm(Count ~ Rank, family=poisson, data=tota, offset=Offset)
summary(fit1)
```

#### Repeat the analysis separately for each Bec Zone, to account for geographic variation

``` {r glm.dens.rank.bybec, echo=T}
totas <- split(tota, tota$BEC)
becglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(becglms), function(i) {
  data.frame(BEC=names(becglms[i]), Intercept=summary(becglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becglms[[i]])$coefficients["Rank","Std. Error"])
}))

kable(sumresults, row.names=F)
rm(sumresults)
```

#### Repeat the analysis separately for each species group, to account for tree productivity differences

``` {r glm.dens.rank.byspp, eval=T, echo=T}
totas <- split(tota, tota$SpeciesGroup)
totas <- totas[-which(names(totas)%in% c(" ", "0", "999", ""))]
spgrpglms <- lapply(totas[c(1:11,13:16, 18:21)], function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(spgrpglms), function(i) {
  data.frame(BEC=names(spgrpglms[i]), Intercept=summary(spgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(spgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(spgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

kable(sumresults, row.names=F)
ggplot(sumresults, aes(x=BEC, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1)

rm(sumresults)
```

#### Repeat the analysis separately for each combination of species group and bec, to account for tree productivity differences

``` {r glm.dens.rank.bybec.byspp, eval=T, echo=T}
tota.sub <- tota[!tota$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota$becspp <- paste(tota$BEC, tota$SpeciesGroup, sep=".")
totas <- split(tota, tota$becspp)
totas <- totas[-c(1,  5,  7, 21, 25, 27, 39, 43, 63, 80, 81, 82)]

becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- lapply(1:length(becspgrpglms), function(i) {
  summary(becspgrpglms[[i]])$coefficients
})

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults <- sumresults[order(sumresults$Rankvs.Density, decreasing=T),]

kable(sumresults, row.names=F)
kable(head(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults)[1] <- "BEC.Species"
sumresults$BEC <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[2]}))

ggplot(sumresults, aes(x=SpeciesGroup, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)


```


**Conclusion:** performance of rank for density dpeends on combination of bec and sp group.

``` {r}
reps <- read.csv("data/Output/BirdSamplingStandTypes-cleaned.2017.04.04.csv", header=T)
becspcountSS <- aggregate(reps$TotalCountSS, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becspcountSS)[ncol(becspcountSS)] <- "CountSS"
becspcountSS$BEC.Species <- paste(becspcountSS$BEC, becspcountSS$SpeciesGroup, sep=".") 

becsparea <- aggregate(reps$ForestArea, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becsparea)[ncol(becsparea)] <- "ForestArea"
becsparea$BEC.Species <- paste(becsparea$BEC, becsparea$SpeciesGroup, sep=".") 

becsp <- merge(becspcountSS, becsparea[c("BEC.Species", "ForestArea")], by="BEC.Species")

sumresults2 <- merge(sumresults, becsp[c("BEC.Species", "ForestArea", "CountSS")], by="BEC.Species", all.x=T)

sumresults2 <- sumresults2[order(sumresults2$CountSS, decreasing=T),]
kable(sumresults2, row.names=F)
```

#### Repeat the analysis including a Rank^2 term. Will maybe give us an idea of why the ranks don't correlate with density.

``` {r glm.dens.rank.rank2.bybec.byspp, eval=F, echo=T}
tota.sub <- tota[!tota$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota$becspp <- paste(tota$BEC, tota$SpeciesGroup, sep=".")
tota$Rank2 <- tota$Rank^2
totas <- split(tota, tota$becspp)


becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank + Rank*Rank, family=poisson, data=x, offset=Offset)
})

removethese <- which(unlist (lapply(1:length(becspgrpglms), function(i) {nrow(summary(becspgrpglms[[i]])$coefficients)}))<3)

totas <- totas[-removethese]

becspgrpglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank + Rank*Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- lapply(1:length(becspgrpglms), function(i) {
  summary(becspgrpglms[[i]])$coefficients
})

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"],
RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"],
Rank2vs.Density=summary(becspgrpglms[[i]])$coefficients["Rank2","Estimate"],
Rank2SE=summary(becspgrpglms[[i]]$coefficients["Rank2", "Std. Error"]))
}))

sumresults <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults <- sumresults[order(sumresults$Rankvs.Density, decreasing=T),]

kable(sumresults, row.names=F)
kable(head(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults[order(sumresults$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults)[1] <- "BEC.Species"
sumresults$BEC <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults$BEC.Species), ".", fixed=T), function(x){x[2]}))

ggplot(sumresults, aes(x=SpeciesGroup, y=Rankvs.Density)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)
```


**Conclusion:** performance of rank for density dpeends on combination of bec and sp group.

``` {r}
reps <- read.csv("data/Output/BirdSamplingStandTypes-cleaned.2017.04.04.csv", header=T)
becspcountSS <- aggregate(reps$TotalCountSS, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becspcountSS)[ncol(becspcountSS)] <- "CountSS"
becspcountSS$BEC.Species <- paste(becspcountSS$BEC, becspcountSS$SpeciesGroup, sep=".") 

becsparea <- aggregate(reps$ForestArea, by=list(BEC=reps$BEC, SpeciesGroup=reps$SpeciesGroup), FUN=function(x){sum(x, na.rm=T)})
colnames(becsparea)[ncol(becsparea)] <- "ForestArea"
becsparea$BEC.Species <- paste(becsparea$BEC, becsparea$SpeciesGroup, sep=".") 

becsp <- merge(becspcountSS, becsparea[c("BEC.Species", "ForestArea")], by="BEC.Species")

sumresults2 <- merge(sumresults, becsp[c("BEC.Species", "ForestArea", "CountSS")], by="BEC.Species", all.x=T)

sumresults2 <- sumresults2[order(sumresults2$CountSS, decreasing=T),]
kable(sumresults2, row.names=F)
```


## Try re-running glm after omitting youngest age class
``` {r}
tota_older <- tota[!tota$ForAge %in% "3_30",]

tota.sub <- tota_older[!tota_older$SpeciesGroup %in% c("Pa", "PyFd_Decid"),]
tota.sub$becspp <- paste(tota.sub$BEC, tota.sub$SpeciesGroup, sep=".")
totasubs <- split(tota.sub, tota.sub$becspp)

becspgrpglms <- lapply(totasubs, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

removethese <- which(unlist (lapply(1:length(becspgrpglms), function(i) {nrow(summary(becspgrpglms[[i]])$coefficients)}))==1)

totasubs <- totasubs[-removethese]

becspgrpglms <- lapply(totasubs, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults_noyoung <- do.call(rbind, lapply(1:length(becspgrpglms), function(i) {
  data.frame(BEC=names(becspgrpglms[i]), Intercept=summary(becspgrpglms[[i]])$coefficients["(Intercept)","Estimate"], Rankvs.Density=summary(becspgrpglms[[i]])$coefficients["Rank","Estimate"], RankSE=summary(becspgrpglms[[i]])$coefficients["Rank","Std. Error"])
}))

sumresults_noyoung <- sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),]

kable(sumresults_noyoung, row.names=F)
kable(head(sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),],10), row.names=F)
kable(tail(sumresults_noyoung[order(sumresults_noyoung$Rankvs.Density, decreasing=T),],10), row.names=F)

names(sumresults_noyoung)[1] <- "BEC.Species"
sumresults_noyoung$BEC <- unlist(lapply(strsplit(as.character(sumresults_noyoung$BEC.Species), ".", fixed=T), function(x){x[1]}))
sumresults_noyoung$SpeciesGroup <- unlist(lapply(strsplit(as.character(sumresults_noyoung$BEC.Species), ".", fixed=T), function(x){x[2]}))
colnames(sumresults_noyoung)[which(colnames(sumresults_noyoung)=="Rankvs.Density")] <- "Rankvs.Density_noyoung"

ggplot(sumresults_noyoung, aes(x=SpeciesGroup, y=Rankvs.Density_noyoung)) + geom_point() + geom_hline(yintercept=0, col="red") + myfacettheme + ylim(-1,1) + facet_wrap(~BEC)

test <- merge(sumresults, sumresults_noyoung[c("BEC.Species","Rankvs.Density_noyoung")], by="BEC.Species")
kable(test[c("BEC.Species", "Rankvs.Density","Rankvs.Density_noyoung")], row.names=F)
```






### GLM of Density ~ Rank

Treat rank as a factor to see if those results are different.

``` {r eval=T, echo=T}
tota$Rank <- as.factor(tota$Rank)
fit1 <- glm(Count ~ Rank, family=poisson, data=tota, offset=Offset)
summary(fit1)

```

Looks like density for each rank is slightly higher than density for rank 1, with exception of rank 6...which has roughly the same density as rank1.

#### Repeat separately for each Bec Zone
``` {r eval=T, echo=T}
totas <- split(tota, tota$BEC)
becglms <- lapply(totas, function(x) {
  fitx <- glm(Count ~ Rank, family=poisson, data=x, offset=Offset)
})

sumresults <- do.call(rbind, lapply(1:length(becglms), function(i) {
  data.frame(BEC=names(becglms[i]), Intercept=summary(becglms[[i]])$coefficients["(Intercept)","Estimate"], Rank2=summary(becglms[[i]])$coefficients["Rank2","Estimate"], Rank3=summary(becglms[[i]])$coefficients["Rank3","Estimate"],
Rank4=summary(becglms[[i]])$coefficients["Rank4","Estimate"],
Rank5=summary(becglms[[i]])$coefficients["Rank5","Estimate"])
}))

kable(sumresults, row.names=F)
```
