---
title: "Bird Abundance Model (For Prediction)" 
author: "Nicole Barker"
date: "Last run: Jan 9, 2018"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

##Script Abstract



## Goal: Build a BRT model that will predict density as a function of VRI attributes, so I can fill in predicted bird density across the BC interior 



**Steps **

1. Load data (Canfor + Forsite)
2. Set training and testing datasets
3. Run BRT with full possible set of covariates
4. Simplify BRT to reduce covariates
5. Generate predictions from BRT
6. In another script .... explore how to convert density to scores and generate my own rank matrix. 


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(
  fig.path = "figures/"
)
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```

## 1. Load data (Canfor + Forsite)

``` {r}
load("cache/bird_vri_dat_ready.RData")
colnames(bird_vri_dat_ready)
unique(bird_vri_dat_ready$VRI_Layer)
tota <- bird_vri_dat_ready

colnames(tota)[which(colnames(tota) == "SS_derived")] <- "SS" # I derived SS from PKEY in a previous file. Here I rename it to SS
colnames(tota)[which(colnames(tota) == "PCODE_derived")] <- "PCODE" # I derived PCODE from PKEY in a previous file. Here I rename it to PCODE

```

### Pick which offset to use

``` {r}
offset_use <- "offset_simple"
#offset_use <- "OFFSET"
```

### Filter to include only one survey per sampling location (SS)

A rigorous approach would involve repeating model building with multiple random sub-samples. However, in the interest of speed, I'm going to extract a single year of data from each SS that has multiple years of data. 

I'll choose the specific year based on its proximity to the VRI reference year. This is just a matter of sorting based on (difference between reference year and bird survey year) and keeping the first one. 

``` {r}
tota <- tota[order(tota$SS, tota$Difference_Bird_Photo),]

dupSS <- tota$SS[duplicated(tota$SS)] # the duplicated SSs

tota.nodup <- tota[!tota$SS %in% dupSS,]
tota.dupl <- tota[tota$SS %in% dupSS,]
rm(tota)

kable(digits=3, tota.dupl[1:25,c("SS", "PKEY", "Difference_Bird_Photo", "Age")])

tota.dupl <- tota.dupl[!duplicated(tota.dupl$SS),] # cuts out all subsequently duplicated rows while keeping the first for each SS

tota <- rbind(tota.nodup, tota.dupl)
colnames(tota)[which(colnames(tota) == "sumABUND")] <- "Count"
```

**NOTES**

* This keeps a total of `r length(unique(tota$SS))` locations, each with `r length(unique(tota$PKEY)) / length(unique(tota$SS))` PKEY. (should be 1. I haven't removed all duplicates if this number is > 1)

## 2. Set training and testing datasets

How many random model replicates (bootstraps) do I want to run? 

``` {r}
boots <- 2
```

``` {r}
SS <- tota$SS
N <- length(SS)
trnsz <- round(N * 0.7)  # Size of the training dataset in # segs
randomsets_tota <- lapply(1:boots, function(i) {
  trndat <- sample(SS, trnsz, replace=F)  # the training datasets
  tstdat <- SS[! SS %in% trndat]
  md.out <- list(train=trndat, test=tstdat)  
  })

names(randomsets_tota) <- paste("Model", c(1:boots), "data", sep=".")

cache("randomsets_tota")
```


## 3. Run BRT with full possible set of covariates

### Set BRT parameters

``` {r}
  tc = 4            # tree complexity 
  bf = 0.5          # bag fraction
  lr = 0.007
  kable(data.frame(BRT.Parameters=rbind(tc, bf, lr)))
```

### Specify predictor variables

``` {r}
pred.variables <- c("BEC_ZONE_CODE", "PROJ_AGE_1", "PROJ_HEIGHT_1", "SPECIES_CD_1", "SPECIES_CD_2", "SPECIES_PCT_1", "SPECIES_PCT_2", "BEC_SUBZONE", "CROWN_CLOSURE", "DistYear_Log", "DistYear_Nonlog") # mark points out that i might want to do years since disturbance.

tota$DistYear_Log <- as.numeric(tota$DistYear_Log)
tota$DistYear_Nonlog <- as.numeric(tota$DistYear_Nonlog)
```

### Load previous results if already run

``` {r}
  if("outboot_tota.RData" %in% dir("../cache") == T) {
    load("cache/outboot_tota.RData")
  }
```

### Run model if not previously run

``` {r}
if(!exists("outboot_tota")) {
outboot_tota <- lapply(1:boots, function(i) {
  run.brt.presample(dat=tota,
                    samples=randomsets_tota[[i]],
                    predictors= pred.variables,
                    measure="Count", 
                    offset=offset_use,
                    tc=tc, bf=bf, lr=lr) #end one BRT
}) #end all BRT boots

cache("outboot_tota")
}
```
  
### summarize results

``` {r, Fig.BRT_output, fig.height=11, fig.width=8.5, dpi=150}
lapply(outboot_tota, function(x) {
  summary(x[["brt"]])
  })

lapply(outboot_tota, function(x) {
  gbm.plot(x[["brt"]], n.plots=6, plot.layout=c(3,2))
  })

```
  
## 4. Simplify BRT to reduce covariates



## 5. Generate predictions from BRT

### Load previous predictions if already run

``` {r}
  
  
```

### Run Prediction Function if not already run

#### Load full VRI dataset (not just intersected with birds)

``` {r eval=F}
load("cache/vri_both.RData")
pred.dat <- vri_both[which(colnames(vri_both) %in% c("ID.Layer.Polygon", pred.variables))]
length(unique(pred.dat$ID.Layer.Polygon))
nrow(pred.dat)

```



``` {r eval=F}
i=1
pred_brt_1 <- predict.brt(tmp.brt=outboot_tota[[i]][["brt"]],
                             pred.dat=pred.dat,
                          id.col="ID.Layer.Polygon")
cache("pred_brt_1")
preds_for_mapping <- pred_brt_1$fullpreds
preds_for_mapping$Layer <- unlist(lapply(strsplit(preds_for_mapping$ID.Layer.Polygon, ".", fixed=T), function(x) {x[1]}))
preds_for_mapping$PolygonID <- unlist(lapply(strsplit(preds_for_mapping$ID.Layer.Polygon, ".", fixed=T), function(x) {x[2]}))

preds_for_mapping_canfor <- subset(preds_for_mapping, subset=preds_for_mapping$Layer == "Canfor_TFL")
preds_for_mapping_forsite <- subset(preds_for_mapping, subset=preds_for_mapping$Layer == "Forsite_VRI")

colnames(preds_for_mapping_canfor)[which(colnames(preds_for_mapping_canfor) == "PolygonID")] <- "OBJECTID"
colnames(preds_for_mapping_forsite)[which(colnames(preds_for_mapping_forsite) == "PolygonID")] <- "FEATURE_ID"

write.table(preds_for_mapping_canfor[c("OBJECTID", "preds")], file="output/predsCanfor_2018.01.09.csv", sep=",", row.names=F, col.names=T)
write.table(preds_for_mapping_forsite[c("FEATURE_ID", "preds")], file="output/predsForsite_2018.01.09.csv", sep=",", row.names=F, col.names=T)

    #predboot_tota <- lapply(function(1:boots), function(i) {
        pred <- predict.brt(tmp.brt=outboot_tota[[i]][["brt"]],
                             pred.dat=pred.dat)
        pred        
      }) 
    #names(predboot_tota) <- names(outboot_tota)
    #rm(outboot_tota)
    #cache(predboot_tota)
  #boots <- length(predboot_MlAbspp_lvl1.1[[1]])
```

