---
title: "Derive matrix from density" 
author: "Nicole Barker"
date: "Last run: Jan 9, 2018"
output: 
  word_document:
    reference_docx: ../styles/ReportFormat_1.docx
---

##Script Abstract



## Goal: Create a matrix in Kari's format that assigns 'ranks' to stands 



**Steps **

1. Load data (Canfor + Forsite)
2. Set training and testing datasets
3. Run BRT with full possible set of covariates
4. Simplify BRT to reduce covariates
5. Generate predictions from BRT
6. In another script .... explore how to convert density to scores and generate my own rank matrix. 


``` {r setup, echo=F, message=F, warning=F}
require(knitr)
opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(
  fig.path = "figures/"
)
```

``` {r load.project, message=F}
require(ProjectTemplate)
load.project()
options(digits=12)
```

## 1. Load predictions

``` {r}
preds_forsite <- read.csv("output/predsForsite_2018.01.09.csv", header=T)
```

## 2. Merge with vri data to obtain forest stand attributes

``` {r}
load("cache/vri_forsite_postclean.RData")

```

#### Clean up

##### Keep only Becs, SpGroups, Age, and Height in the mastermatrix

``` {r}
vri_forsite_postclean$ForID <- with(vri_forsite_postclean, paste(BecZone, SpeciesGroup, Age, Height, sep="."))
vri_forsite_Inmastermatrix <- subset(vri_forsite_postclean, vri_forsite_postclean$ForID %in% mastermatrix$ForID)
```

#### Identify desired columns

``` {r}
cols_to_keep_prediction_matrix <- c("FEATURE_ID", "BecZone", "BEC_SUBZONE", 
                                    "Age", "Height", "SpeciesGroup", 
                                    "SPECIES_CD_1", "SPECIES_CD_2", "Rank") 

cols_to_keep_kari_matrix <- c("BecZone", "Age", "Height", "SpeciesGroup", "preds") 
```

#### Create dataset with VRI attributes and predictions

``` {r}
preds_forsite_small <- merge(preds_forsite, vri_forsite_Inmastermatrix[cols_to_keep_prediction_matrix], by="FEATURE_ID")
```

## 3. Aggregate into various matrix formats

### Kari's matrix

``` {r}
preds_forsite_kariMatrix <- preds_forsite_small[cols_to_keep_kari_matrix]

colnames(preds_forsite_kariMatrix)[which(colnames(preds_forsite_kariMatrix) == "preds")] <- "value"

kable(preds_forsite_kariMatrix[1:15,], row.names=F)

lapply(preds_forsite_kariMatrix[1:4], function(x) {unique(x)})
```

#### Cast out 

``` {r}
pred_matrix_Kariformat <- cast(preds_forsite_kariMatrix, BecZone + SpeciesGroup ~ Age + Height , fun.aggregate = mean) 

pred_matrix_Kariformat <- pred_matrix_Kariformat[c("BecZone", "SpeciesGroup", 
                                                   "3_30_1", "3_30_2", "3_30_3", "3_30_4+", 
                                                   "31_80_1", "31_80_2", "31_80_3", "31_80_4+", 
                                                   "81_120_1", "81_120_2", "81_120_3", "81_120_4+",
                                                   "121_250_1", "121_250_2", "121_250_3", "121_250_4+",
                                                   "250_1", "250_2", "250_3", "250_4+")]

write.table(pred_matrix_Kariformat, file="output/Matrix_Density_KariFormat.csv", sep=",", row.names=F, col.names=T)

```

#### Scale to 6

``` {r}

tmp <- pred_matrix_Kariformat[3:ncol(pred_matrix_Kariformat)]

pred_matrix_Kariformat_scaled <- cbind(pred_matrix_Kariformat[1:2], 6*(tmp / max(tmp, na.rm=T)))

summary(pred_matrix_Kariformat_scaled)

write.table(pred_matrix_Kariformat_scaled, file="output/Matrix_Density_KariFormat_scaled.csv", sep=",", row.names=F, col.names=T)

```
